A quantitative framework for investigating the reliability of network construction

### 2018-03-08 manuscript/CI_interactions.tex
spellchecked SI
intro looks good to me

    Representing an assemblage of species as a network offers a convenient summary of how the community is constructed as it simultaneously accounts for species composition and interactions between species. A tabulation of the nodes (species) and their relative abundances forms the basis for traditional metrics of community composition. To move from these simpler metrics to a network framework, the tabulation of nodes is combined with interactions (links between nodes) so that networks provide additional, higher-order information on community structure. While this additional information is useful (as, for example, interactions can affect changes in species abundances over time), empirical descriptions of ecological networks are still limited because they are usually considered to be static representations of the communities and interactions they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

    The assumption of static communities contrasts significantly with the widely recognised stochasticity of ecological communities~\citep{Gotelli2000}. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. Likewise, interactions vary over space~\citep{Kitching1987,Baiser2012}, time~\citep{Kitching1987,Olesen2011a}, and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}. Moreover, variability in community composition and interactions may or may not be closely related. The removal of a species from a site will obviously also remove its interactions but, conversely, the co-occurrence of potentially interacting species does not in itself guarantee that they will interact at a given place and time. Interactions can be lost if the interaction partners remain present but are separated in time or are too rare to detect each other~\citep{Tylianakis2010}. Interactions can also fail to occur because of environmental contingencies~\citep{Poisot2015}, or through changes to individual preferences~\citep{Fodrie2015}.

@@

    In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions [[double-check that we do this]]. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

    [[check that all the figs are with Zillis]]

@@

      In a strict Bayesian framework, we wish to use a prior distribution that does not rely on any information from the study at hand. Network data for a similar study system may, however, not be available. In that case, one might use the first sub-network collected as ``training data" to guide future sampling. To simulate this situation, we created priors using a single sub-network from the middle of the geographical distribution of the~\citep{Kopelke2017} dataset (Zillis sub-network). To demonstrate how the use of a data from a different system can affect the prior distribution and conclusions based on it, we repeated our analyses using priors derived from a much smaller \emph{Salix}-galler-natural enemy system~\citep{Barbour2016,Barbour2016Dryad}. This smaller system was much more densely-connected than that described in~\citet{Kopelke2017} and provided unreasonable distributions for interaction probabilities (\emph{Appendix SX}).

@@

      For species where no co-occurrences were observed ($n=0$), we can calculate the estimates for the mean and variance of $\lambda_{ij}$ directly from the prior parameters following equations~\ref{mean} and~\ref{variance} (see \emph{Appendix S1} for R implementation). For the \emph{Salix}-galler network, the prior distribution  was: $\bar\lambda$=0.028, var($\lambda$)=8.60$\times$10$^{-5}$. The prior distribution for the galler-natural enemy network was: $\bar\lambda$=0.076, var($\lambda$)=0.008. The posterior interaction probabilities obtained based on the Zillis sub-network were much lower than those obtained based on~\citet{Barbour2016}; this emphasises the importance of using an appropriate study system when constructing a prior (\emph{Appendix SX}).

@@

      \begin{figure}[ht]
        \caption{The number of samples required to achieve a given level of confidence that an interaction probability $\lambda_{ij}$ is below a given threshold varies with both parameters. With a low threshold, our confidence that $\lambda_{ij}$ is below the threshold increases rapidly with repeated observation of co-occurrence without interaction. Here we show the cumulative density functions for threshold probabilities of 0.5 (solid line), 0.25 (dashed line), 0.1 (dash-dot line), and 0.05 (dotted line) as well as the points at which the cdf reaches 0.90 (orange square), 0.95 (red circle), and 0.975 (blue diamond) for each threshold value. The large ticks along the x-axis indicate the number of samples associated with each of these points. \textbf{A)} In the \emph{Salix}-galler network component, the 95\% credible interval for $\lambda$ when $n$=0 was (0.013, 0.049). We can therefore be at least 95\% confident that $\lambda$ is below thresholds of 0.1 or 0.05 without any observed co-occurrence. To be confident that $\lambda$ is less than 0.01, however, would require more observed co-occurrences than there are sites in our dataset. \textbf{B)} In the galler-parasitoid network component, the 95\% credible interval for $\lambda$ was substantially broader and many observed co-occurrences (\approx 15-35) are required to be 95\% confident that $\lambda$ is below thresholds of 0.1 or 0.05.}
        \label{Salix_cdfs}
        \includegraphics[width=.8\textwidth]{figures/Salix_Galler_samples_and_cdfs_Zillis.eps}

>>>>

    Representing an assemblage of species as a network offers a convenient summary of how the community is constructed as a network simultaneously describes species composition and interactions between species. A tabulation of the nodes (species) and their relative abundances forms the basis for traditional metrics of community composition. To move from these simpler metrics to a network framework, the tabulation of nodes is combined with interactions (links between nodes) so that networks provide additional, higher-order information on community structure. While this additional information is useful (as, for example, interactions can affect changes in species abundances over time), empirical descriptions of ecological networks are still limited because they are usually considered to be static representations of the communities and interactions they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}.

    The assumption of static communities contrasts significantly with the widely recognised stochasticity of ecological communities~\citep{Gotelli2000}. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. Likewise, interactions vary over space~\citep{Kitching1987,Baiser2012}, time~\citep{Kitching1987,Olesen2011a}, and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}. We emphasise that variability in community composition and interactions may or may not be closely related. The removal of a species from a site will obviously also remove its interactions but, conversely, the co-occurrence of potentially interacting species does not in itself guarantee that they will interact at a given place and time. Interactions can be lost if the interaction partners remain present but are separated in time or are too rare to detect each other~\citep{Tylianakis2010}. Interactions can also fail to occur because of environmental contingencies~\citep{Poisot2015}, or through changes to individual preferences~\citep{Fodrie2015}.

@@

    In this study, we formalize the description of interactions between species as probabilities and develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions [[double-check that we do this]]. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

@@

      In a strict Bayesian framework, we wish to use a prior distribution that does not rely on any information from the study at hand. Network data for a similar study system may, however, not be available. In that case, one might use the first sub-network collected as ``training data" to guide future sampling. To simulate this situation, we created priors using a single sub-network from the middle of the geographical distribution of the~\citep{Kopelke2017} dataset (Zillis sub-network). To demonstrate how the use of a data from a different system can affect the prior distribution and conclusions based on it, we repeated our analyses using priors derived from a much smaller \emph{Salix}-galler-natural enemy system~\citep{Barbour2016,Barbour2016Dryad}. This smaller system was much more densely-connected than that described in~\citet{Kopelke2017} and provided unreasonable distributions for interaction probabilities (\emph{Appendix S4}).

@@

      For species where no co-occurrences were observed ($n=0$), we can calculate the estimates for the mean and variance of $\lambda_{ij}$ directly from the prior parameters following equations~\ref{mean} and~\ref{variance} (see \emph{Appendix S1} for R implementation). For the \emph{Salix}-galler network, the prior distribution  was: $\bar\lambda$=0.028, var($\lambda$)=8.60$\times$10$^{-5}$. The prior distribution for the galler-natural enemy network was: $\bar\lambda$=0.076, var($\lambda$)=0.008. The posterior interaction probabilities obtained based on the Zillis sub-network were much lower than those obtained based on~\citet{Barbour2016}; this emphasises the importance of using an appropriate study system when constructing a prior (\emph{Appendix S4}).

@@

      \begin{figure}[ht]
        \caption{The number of samples required to achieve a given level of confidence that an interaction probability $\lambda_{ij}$ is below a given threshold varies with both parameters. With a low threshold, our confidence that $\lambda_{ij}$ is below the threshold increases rapidly with repeated observation of co-occurrence without interaction. Here we show the cumulative density functions for threshold probabilities of 0.5 (solid line), 0.25 (dashed line), 0.1 (dash-dot line), and 0.05 (dotted line) as well as the points at which the cdf reaches 0.90 (orange square), 0.95 (red circle), and 0.975 (blue diamond) for each threshold value. The large ticks along the x-axis indicate the number of samples associated with each of these points. \textbf{A)} In the \emph{Salix}-galler network component, the 95\% credible interval for $\lambda_{ij}$ when $n$=0 was (0.013, 0.049). We can therefore be at least 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05 without any observed co-occurrence of species $i$ and $j$. To be confident that $\lambda_{ij}$ is less than 0.01, however, would require more observed co-occurrences than there are sites in our dataset. \textbf{B)} In the galler-parasitoid network component, the 95\% credible interval for $\lambda_{ij}$ was substantially broader and many observed co-occurrences ($\approx$ 15-35) are required to be 95\% confident that $\lambda_{ij}$ is below thresholds of 0.1 or 0.05.}
        \label{Salix_cdfs}
        \includegraphics[width=.8\textwidth]{figures/Salix_Galler_samples_and_cdfs_Zillis.eps}
